---
title: "Data Exploration And Visualization"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(caret)
library(dplyr)
library(MLmetrics)
library(naivebayes)
library(randomForest)
library(reshape)
library(kernlab)
# devtools::install_version('DMwR', '0.4.1')
library("DMwR")
set.seed(12)
```

## Read the data

```{r}
data <- read.csv('./data/PS_20230404203710_1288618611/PS_20230404203710_1288618611_log.csv', nrows=10000)
```

## Exploratory analysis

#### Graph of what action are tagged as isFraud

```{r}
ggplot(data, aes(x=type, y=isFraud, fill=type)) + 
  geom_bar(stat = "identity")
```
#### Histogram of amounts tagged as frauds

```{r}
ggplot(data[data$isFraud==1,], aes(x=amount)) + 
  geom_histogram(bins = 20)
```
#### Top recepients accounts tagged as fraud based on amount

```{r}
fraud_data <- data[data$isFraud==1,]
ordered_data <- fraud_data[order(fraud_data$amount),]

ggplot(tail(ordered_data, 10), aes(x=nameDest, y=amount, fill=nameDest)) + 
  coord_flip() +
  geom_bar(stat = "identity", ) 
```

## Preprocessing

### Convert the output as factor

```{r}
data$isFraud <- as.factor(data$isFraud)
data$type <- as.factor(data$type)
```

### Select required columns based on variances

```{r}
# Removing vars with 0 variance
sdata <- select(data, type, amount, oldbalanceOrg, newbalanceOrig, oldbalanceDest, newbalanceDest, isFraud)
``` 

### Check and fix class imbalance so that we can apply our algorithms

```{r}
table(sdata$isFraud)

# Class is highly imbalanced, we can fix it
sdata <- SMOTE(isFraud ~ ., sdata, perc.over = 2000, perc.under = 400)

# Verify if the balance is okay
table(sdata$isFraud)
```

### Split the data into training and test

```{r}
split_index <- createDataPartition(y = sdata$isFraud, times = 1, p = 0.6, list = FALSE)
training_data <- sdata[split_index,]
test_data <- sdata[-split_index,]
dim(training_data)
dim(test_data)
```

### Create control for Cross validation

```{r}
control <- trainControl(method = "repeatedcv", number = 2, repeats = 2)
```

## Training & Evaluation of Models

### K nearest neighbours

#### Training 

```{r}
knn.model <- train(isFraud ~ ., 
                   method='knn', 
                   data = training_data, 
                   preProcess = c("center", "scale"), 
                   trControl=control)

plot(knn.model)
```

#### Validation

```{r}
predictions <- predict(knn.model, test_data)
confusionMatrix(predictions, test_data$isFraud )
F1_Score(test_data$isFraud, predictions)
```
### Random Forest

#### Training 

```{r}
rf.model <- train(isFraud ~ ., 
                   method='rf', 
                   data = training_data, 
                   preProcess = c("center", "scale"), 
                   trControl=control)

plot(rf.model)
```

#### Validation

```{r}
predictions <- predict(rf.model, test_data)
confusionMatrix(predictions, test_data$isFraud )
F1_Score(test_data$isFraud, predictions)
```


### Naive Bayes

#### Training 

```{r}
nb.model <- train(isFraud ~ ., 
                   method='naive_bayes', 
                   data = training_data, 
                   preProcess = c("center", "scale"), 
                   trControl=control)

plot(nb.model)
```

#### Validation

```{r}
predictions <- predict(nb.model, test_data)
confusionMatrix(predictions, test_data$isFraud )
F1_Score(test_data$isFraud, predictions)
```


### Support vector machines

#### Training 

```{r}
svm.model <- train(isFraud ~ ., 
                   method='svmLinear', 
                   data = training_data, 
                   preProcess = c("center", "scale"), 
                   trControl=control)

```

#### Validation

```{r}
predictions <- predict(svm.model, test_data)
confusionMatrix(predictions, test_data$isFraud )
F1_Score(test_data$isFraud, predictions)
```

### Neural Networks

#### Training 

```{r}
nn.model <- train(isFraud ~ ., 
                   method='nnet', 
                   data = training_data, 
                   preProcess = c("center", "scale"), 
                   trControl=control)

plot(nn.model)
```

#### Validation

```{r}
predictions <- predict(nn.model, test_data)
confusionMatrix(predictions, test_data$isFraud )
F1_Score(test_data$isFraud, predictions)
```